BOAST: a Metaprogramming Framework to Produce Portable and Efficient Computing Kernels for HPC Applications

### 2016-01-12 Paper/BOAST.tex
WIP: integrating Fred remarks.

%% or include affiliations in footnotes:
\author[mymainaddress]{Brice Videau}
\ead{brice.videau@imag.fr}
\author{Kevin Pouget}
\author{Luigi Genovese}
\author{Thierry Deutsch}
\author{Dimitri Komatitsch}
\author{Jean-François Méhaut}

\address[mymainaddress]{LIG/CNRS}

\begin{abstract}
The Abstract.
\end{abstract}

\begin{keyword}
Code Generation \sep Portability \sep Genericity \sep Productivity and Software
Design \sep High Performance Computing \sep Autotuning \sep Non Regression
Testing
\end{keyword}

@@

Porting and tuning HPC applications to new platforms is tedious and costly in
terms of human resources. Nonetheless, it is a very important aspect of the
Mont-Blanc project. Indeed, for the project, more than ten applications were
selected to be ported and optimized for the prototype platform.

Unfortunately, portability efforts are often lost when migrating to a new
architecture. Worse, code may lose maintainability because several versions of
some functionalities coexist, usually with a lot of duplication.

Thus productivity of porting and tuning efforts is low as a huge fraction of
those developments are never used after the platform they were intended for is
decommissioned.

Genericity of HPC codes is often limited. One of the reason is that producing
generic code in Fortran 90/95 is difficult as the language does not really fit
for it. Sometimes, adding genericity degrades performance as optimization
opportunities that come from over-specification are lost.

Functionality of HPC codes is tied to the previous point. Without genericity,
adding new functionalities can be quite costly.

@@

@@

Evolution of HPC Architectures is rapid and also diverse: in the last 5
years no less than 6 architectures have been number one in the Top500:
\begin{itemize}
\item Intel Processor + Xeon Phi (Tianhe-2)
\item AMD Processor + NVIDIA GPU (Titan)
\item IBM BlueGene/Q (Sequoia)
\item Fujitsu SPARC64 (K computer)
\item Intel Processor + NVIDIA GPU (Tianhe-1)
\item AMD Processor (Jaguar)
\end{itemize}
Being able to efficiently use those architectures on such a small
time-frame is challenging.

The race to exascale is not going to simplify the environment. All of the above
architectures can be considered. Network architectures also can be very
diverse.  For instance European FP7 project DEEP considers using Accelerators
(XEON Phi) while the European FP7 project Mont-Blanc considers using low-power
embedded processor with integrated GPU.

Running existing applications on those new architectures is an open research
subject as well as an ongoing porting effort.  Thus, those projects have work
packages dedicated to applications. Those work packages are dedicated to porting
and optimizing Scientific applications on those new architectures. In the DEEP
project six applications were selected for porting and optimizing, eleven were
selected in the Mont-Blanc project.

@@

Scientific Computing Applications are usually developed by physicist, chemist or
meteorologist. Those codes are usually written in Fortran for historical and
performance reasons. Codes can be quite huge (several thousands lines of code [LOC])
with lots of functionalities. Nonetheless, they are usually based on computing
kernels.  Computing kernels are resource intensive and well-defined parts of a
program that usually work on precisely defined data. Those kernels represent the most
time-consuming part of an HPC application, and consequently, they are the prime target for
optimization.

Those applications are often developed by several individuals. Sometimes, some
of those developers only work a few months on the application. Maintaining
optimized code written by someone else is quite a challenge. Several languages
and programming paradigms can also be used in a project and thus the maintainer
must be knowledgeable in several areas of expertise. Portability problems can
also be caused by the availability of an optimizing compiler for a specific
language and architecture. C compilers are usually the first available while
Fortran may sometimes arrive later.

@@

The problematic here is to obtain computing kernels that present good
performances on the architectures encountered by the application while still
being portable after the optimization process took place. Indeed the
application might encounter one of the many architectures that can be found in
HPC. Investing manpower to optimize the application for a new architecture is
reasonable, suffering hindrance from previous optimization work is not. Thus
optimizations have to be as orthogonal as possible from one another so as to be
easily activated and deactivated.

@@

What we want is computing kernels that are written:
\begin{itemize}
\item in a \emph{portable} manner,
\item in a way that raises developer \emph{productivity},
\item and, present good \emph{performance}.
\end{itemize}

\section{Motivating Example: OpenCL Laplace}
\label{sec:laplace}

During one of the Mont-Blanc face to face meeting a talk on OpenCL optimization
on the Mali GPU was given. One of the case study was a Laplace filter on the
GPU~\cite{opencl_arm_training}.

@@

\lstset{style=customc,
        caption={Laplace C99 Filter},
        breaklines=true,
        breakautoindent=false,
        label={lst:laplace_c_filter}}
\begin{lstlisting}
void laplace(const int width,
             const int height,
             const unsigned char src[height][width][3],
                   unsigned char dst[height][width][3]){
  for (int j = 1; j < height-1; j++) {
    for (int i = 1; i < width-1; i++) {
      for (int c = 0; c < 3; c++) {
        int tmp = -src[j-1][i-1][c] -   src[j-1][i][c] - src[j-1][i+1][c]\
                 - src[j  ][i-1][c] + 9*src[j  ][i][c] - src[j  ][i+1][c]\
                 - src[j+1][i-1][c] -   src[j+1][i][c] - src[j+1][i+1][c];
        dst[j][i][c] = (tmp < 0 ? 0 : (tmp > 255 ? 255 : tmp));
      }
    }
  }
}
\end{lstlisting}

@@

\lstset{style=customcl,
        caption={Laplace OpenCL Filter},
        breaklines=true,
        breakautoindent=false,
        label={lst:laplace_opencl_filter}}
\begin{lstlisting}
kernel laplace(const int width,
               const int height,
               global const uchar *src,
               global       uchar *dst){
  int i = get_global_id(0);
  int j = get_global_id(1);
  for (int c = 0; c < 3; c++) {
    int tmp = -src[3*width*(j-1) + 3*(i-1) + c]\
             - src[3*width*(j-1) + 3*(i  ) + c]\
             - src[3*width*(j-1) + 3*(i+1) + c]\
             - src[3*width*(j  ) + 3*(i-1) + c]\
           + 9*src[3*width*(j  ) + 3*(i  ) + c]\
             - src[3*width*(j  ) + 3*(i+1) + c]\
             - src[3*width*(j+1) + 3*(i-1) + c]\
             - src[3*width*(j+1) + 3*(i  ) + c]\
             - src[3*width*(j+1) + 3*(i+1) + c];
    dst[3*width*j + 3*i + c] = clamp(tmp, 0, 255);
  }
}
\end{lstlisting}

@@

\paragraph{Vectorization}
The first proposed optimization involves computing five pixels instead of one using the
vectors available to the Mali architecture. The vectors have to reside in memory to
ensure an efficient execution. Using vectors of 16 elements, 15 useful components
are simultaneously loaded and can be used in computation. This optimization
yields speedups between 1.5 and 6 depending on the image size.

\lstset{style=customcl,
        caption={Laplace OpenCL Filter Vectorized},
        breaklines=true,
        breakautoindent=false,
        label={lst:laplace_opencl_filter_vector}}
\begin{lstlisting}
kernel laplace(const int width,
               const int height,
               global const uchar *src,
               global       uchar *dst){
  int i = get_global_id(0);
  int j = get_global_id(1);

@@

\lstset{style=customcl,
        caption={Laplace OpenCL Filter Vectorized with Synthesized Loads},
        breaklines=true,
        breakautoindent=false,
        label={lst:laplace_opencl_filter_vector_synth}}
\begin{lstlisting}
  uchar16 v11_ = vload16( 0, src + 3*width*(j-1) + 3*5*i - 3 );
  uchar16 v13_ = vload16( 0, src + 3*width*(j-1) + 3*5*i + 3 );
  uchar16 v12_ = uchar16( v11_.s3456789a, v13_.s56789abc );
  uchar16 v21_ = vload16( 0, src + 3*width*(j  ) + 3*5*i - 3 );
  uchar16 v23_ = vload16( 0, src + 3*width*(j  ) + 3*5*i + 3 );
  uchar16 v22_ = uchar16( v21_.s3456789a, v23_.s56789abc );
  uchar16 v31_ = vload16( 0, src + 3*width*(j+1) + 3*5*i - 3 );
  uchar16 v33_ = vload16( 0, src + 3*width*(j+1) + 3*5*i + 3 );
  uchar16 v32_ = uchar16( v31_.s3456789a, v33_.s56789abc );
\end{lstlisting}

@@

\lstset{style=customcl,
        caption={Laplace OpenCL Filter Vectorized Using \textit{short}},
        breaklines=true,
        breakautoindent=false,
        label={lst:laplace_opencl_filter_short}}
\begin{lstlisting}
  short16 v11 = convert_short16(v11_);
  short16 v12 = convert_short16(v12_);
  short16 v13 = convert_short16(v13_);
  short16 v21 = convert_short16(v21_);
  short16 v22 = convert_short16(v22_);
  short16 v23 = convert_short16(v23_);
  short16 v31 = convert_short16(v31_);
  short16 v32 = convert_short16(v32_);
  short16 v33 = convert_short16(v33_);

@@

Nonetheless, the process is frustrating. The optimizations are never evaluated
independently from one another. Some were arbitrarily configured (the number of
pixels chosen for instance). Testing the different combinations of those
optimizations and the different parameters would be very costly in developer
time and, moreover, like all repetitive and tedious tasks, error prone. Thus
every created kernel would have to be thoroughly tested to ensure no error was
done.

@@

In order to clearly differentiate what is going to be generated from what is
related to manipulations in the host language four keywords were defined. They
are \emph{decl}, \emph{pr}, \emph{opn} and \emph{close}. As the language is an
EDSL, these four keywords are methods in the BOAST namespace. Sample usage of
these keywords will be found in the next Figures.
%Usage of these keywords can be found on Figure~\ref{fig:keywords}.

@@

\begin{figure}
\lstset{style=BOAST,
        caption={BOAST code}, 
        label={lst:BOAST_algebra},
        numbers=left }
\begin{lstlisting}
i = BOAST::Int("i") # or BOAST::Variable::new("i", Int)
k = BOAST::Int("k", :size => 8)
l = BOAST::Real("l", :dim => [ BOAST::Dim(-5, 21) ], :local => true )
BOAST::decl i, k, l
BOAST::pr i === 5
j = i + 5
BOAST::pr k === j * 2
BOAST::pr l[k] === 1.0
BOAST::register_funccall("sin")
BOAST::pr l[k+1] === BOAST::sin(j)
\end{lstlisting}

@@

\begin{figure}
\lstset{style=BOAST,
        caption={BOAST code}, 
        label={lst:BOAST_control},
        numbers=left }
\begin{lstlisting}
BOAST::register_funccall("modulo")
i = BOAST::Int("i")
j = BOAST::Int("j")
BOAST::pr j === 0
BOAST::pr BOAST::For( i, 0, 100 ) {
  BOAST::pr BOAST::If( BOAST::modulo(i,7) == 0 ) {
    BOAST::pr j === j + 1
  }
}
\end{lstlisting}

@@

\begin{figure}
\lstset{style=BOAST,
        caption={BOAST code}, 
        label={lst:BOAST_procedure},
        numbers=left }
\begin{lstlisting}
n = Int("n", :dir => :in)
ndat = Int("ndat", :dir => :in)
x = Real("x", :dir => :in, :dim => [ Dim(0, n-1), Dim(ndat) ])
y = Real("y", :dir => :out, :dim => [ Dim(ndat), Dim(0, n-1) ])
p = Procedure("magicfilter", [n, ndat, x, y])
opn p
close p
\end{lstlisting}

>>>>

%% or include affiliations in footnotes:
\author[lig,cnrs]{Brice Videau}
\ead{brice.videau@imag.fr}
\author[lig,ujf]{Kevin Pouget}
\author[cea]{Luigi Genovese}
\author[cea]{Thierry Deutsch}
\author[cnrs]{Dimitri Komatitsch}
\author[lig,inria]{Frédéric Desprez}
\author[lig,ujf]{Jean-François Méhaut}

\address[lig]{Laboratoire d'Informatique de Grenoble}
\address[cnrs]{Centre National de la Recherche Scientifique}
\address[cea]{CEA/INAC/SP2M/L\_Sim}
\address[ujf]{Université Joseph Fourier}
\address[inria]{Inria}
\begin{abstract}
The Abstract.
\end{abstract}

\begin{keyword}
Code Generation \sep Portability \sep Genericity \sep Autotuning \sep Non
Regression Testing \sep Productivity and Software Design \sep High Performance
Computing
\end{keyword}

@@

Porting and tuning HPC applications to new platforms is tedious and costly in
term of human resources.  Many developer teams are more concentrated on solving
the physical problems faster and better with new algorithms and simulating new
physics. This is why optimization, while important, is a task which is
sometimes neglected or hidden for many developers.  In the same way, Domain
Specific Language (DSLs) have exactly this goal, the complete separation of the
high-level code dedicated to simulate new physics and the low-level one which
is more dependent on the HPC architecture. Nevertheless, decoupling the high
and low levels don't avoid the need of optimization which becomes more and more
unavoidable with new architectures based on many cores and accelerators.

Unfortunately, portability efforts of the low level i.e. computing kernels are
often lost when migrating to a new architecture. Worse, code may lose
maintainability because several versions of some functionalities coexist,
usually with a lot of duplication.  Thus productivity of porting and tuning
efforts is low as a huge fraction of those developments are never used after
the platform they were intended for is decommissioned.

The primary goal of the BOAST metaprogramming framework is to simplify the
optimization of computing kernels for HPC applications. The main idea is to use
a high level language to define an API of computing kernels i.e.  subroutines
and their related arguments, but also describe the way to optimize it
(unrolling, vectorization, parametrization, templating, tiling).  BOAST will
test automatically the many possibilities of optimizations and choose the best
one after validation from a reference version.  Then BOAST can generate pieces
of Fortran, C, OpenCL, or CUDA codes, test it for a specific target
architecture.

With the definition of an API, BOAST can hide the implementation details at
high level but also is a way to factorize different kernels depending only on
parameters such as data types, length of filters or boundary conditions.  The
maintainability and genericity of HPC codes is often limited. One of the reason
is that producing generic code in Fortran 90/95 is difficult as the language
does not really fit for it. Sometimes, adding genericity or functionalities can
be quite costly and degrades performance as optimization opportunities that
come from over-specification are lost. BOAST can provide a solution to handle
genericity and add new functionalities at low levels.

%For instance, the optimization of short convolutions depends strongly on the
%length of the filters. This is why HPC codes using finite differences handle
%only one order. BOAST allows handling any order and even more generalizing to
%another numerical scheme.

The paper is organized as follows. First we further motivate the need for
metaprogramming frameworks.  Secondly we illustrate the optimization problems
related to scientific computing with the Laplace equation. Thirdly we present
our solution, called BOAST, its architecture and its functionalities. Fourthly
we apply BOAST to three use cases: the Laplace equation motivating example;
short multi-dimensional convolutions from the BigDFT quantum chemistry
application; and the seismic wave propagation simulator SPECFEM3D where 80 GPU
computing kernels have been ported improving, considerably the maintainability
and providing a single code base for OpenCL and CUDA kernel versions. Finally
we compare BOAST with other related works.

@@

The motivation of the BOAST project comes from the need of the HPC communities
which have full and large applications.  These applications are difficult to
port and maintain for emerging architectures. Especially since new
architectures are based on accelerators, longer vector units, or low power
processors.

@@

Evolution of HPC Architectures is rapid and also diverse: in the last 7 years
no less than 6 architectures have been number one in the
Top500~\cite{webtop500}:
\begin{itemize}
\item Intel Processor + Xeon Phi (Tianhe-2, 2013)
\item AMD Processor + NVIDIA GPU (Titan, 2012)
\item IBM BlueGene/Q (Sequoia, 2012)
\item Fujitsu SPARC64 (K computer, 2011)
\item Intel Processor + NVIDIA GPU (Tianhe-1, 2010)
\item AMD Processor (Jaguar, 2009)
\end{itemize}
Being able to efficiently use those architectures on such a small period is
challenging for scientists and application developers.

The race to exascale is not going to simplify the environment. All of the above
machine architectures can be considered to build these supercomputers and other
will appear soon. For instance European FP7 project DEEP considers using
Accelerators (XEON Phi) while the European FP7 project Mont-Blanc considers
using low-power embedded processor with integrated GPU.

Running existing applications on these new architectures is an open research
subject as well as an ongoing porting effort.  Thus, those projects have work
packages dedicated to applications. Those work packages are dedicated to
porting and optimizing Scientific applications on those new architectures. In
the DEEP project six applications were selected for porting and optimizing,
eleven were selected in the Mont-Blanc project.

@@

Scientific Computing Applications are usually developed by other science
researchers such as physicists, chemists or meteorologists. Those codes are
usually written in Fortran for historical and performance reasons. Codes can be
quite huge (several thousands lines of code [LOC]) with lots of
functionalities. Nonetheless, they are usually based on computing kernels.
Computing kernels are resource intensive and well-defined parts of a program
that usually work on precisely defined data. Those kernels represent the most
time-consuming part of an HPC application, and consequently, they are the prime
target for optimization.

Those applications are often developed by several individuals. Sometimes, some
of those developers only work a few months on the application. Maintaining
optimized code written by someone else is quite a challenge. Several languages
and programming paradigms can also be used in a project and thus the maintainer
must be knowledgeable in several areas of expertise. Portability problems can
also be caused by the availability of an optimizing compiler for a specific
language and architecture. C compilers are usually the first available while
Fortran may sometimes come later.

@@

The problematic here is to obtain computing kernels that present good
performance on the architectures encountered by the application while staying
portable after the optimization process took place. Indeed the application
might encounter one of the many architectures that can be found in HPC.
Investing manpower to optimize the application for a new architecture is
reasonable, suffering hindrance from previous optimization work is not. Thus
optimizations have to be as orthogonal as possible from one another so as to be
easily activated and deactivated.

@@

What we want is computing kernels that are written in a \emph{portable} manner,
in a way that raises developer \emph{productivity}, and they must present good
\emph{performance}.

\section{Basic Example: OpenCL Laplace}
\label{sec:laplace}

During one of the Mont-Blanc face to face meeting a talk on OpenCL optimization
on the Mali GPU was given. One of the case study was a Laplace filter~\cite{opencl_arm_training}.

@@

\lstset{style=customc,
        caption={Laplace C99 Filter.},
        breaklines=true,
        breakautoindent=false,
        label={lst:laplace_c_filter}}
\begin{lstlisting}
void laplace(const int width,
             const int height,
             const unsigned char src[height][width][3],
                   unsigned char dst[height][width][3]){
  for (int j = 1; j < height-1; j++) {
    for (int i = 1; i < width-1; i++) {
      for (int c = 0; c < 3; c++) {
        int tmp = -src[j-1][i-1][c] -   src[j-1][i][c] - src[j-1][i+1][c]\
                 - src[j  ][i-1][c] + 9*src[j  ][i][c] - src[j  ][i+1][c]\
                 - src[j+1][i-1][c] -   src[j+1][i][c] - src[j+1][i+1][c];
        dst[j][i][c] = (tmp < 0 ? 0 : (tmp > 255 ? 255 : tmp));
      }
    }
  }
}
\end{lstlisting}

@@

\lstset{style=customcl,
        caption={Laplace OpenCL Filter.},
        breaklines=true,
        breakautoindent=false,
        label={lst:laplace_opencl_filter}}
\begin{lstlisting}
kernel laplace(const int width,
               const int height,
               global const uchar *src,
               global       uchar *dst){
  int i = get_global_id(0);
  int j = get_global_id(1);
  for (int c = 0; c < 3; c++) {
    int tmp = -src[3*width*(j-1) + 3*(i-1) + c]\
             - src[3*width*(j-1) + 3*(i  ) + c]\
             - src[3*width*(j-1) + 3*(i+1) + c]\
             - src[3*width*(j  ) + 3*(i-1) + c]\
           + 9*src[3*width*(j  ) + 3*(i  ) + c]\
             - src[3*width*(j  ) + 3*(i+1) + c]\
             - src[3*width*(j+1) + 3*(i-1) + c]\
             - src[3*width*(j+1) + 3*(i  ) + c]\
             - src[3*width*(j+1) + 3*(i+1) + c];
    dst[3*width*j + 3*i + c] = clamp(tmp, 0, 255);
  }
}
\end{lstlisting}

@@

\paragraph{Vectorization}
The first proposed optimization involves computing five pixels instead of one
using the vectors available on the Mali architecture. The vectors have to be
kept in memory to ensure an efficient execution. Using vectors of 16 elements,
15 useful components are simultaneously loaded and can be used in computation.
This optimization yields speedups between 1.5 and 6 depending on the image
size.

\lstset{style=customcl,
        caption={Laplace OpenCL Filter Vectorized.},
        breaklines=true,
        breakautoindent=false,
        label={lst:laplace_opencl_filter_vector}}
\begin{lstlisting}
kernel laplace(const int width,
               const int height,
               global const uchar *src,
               global       uchar *dst){
  int i = get_global_id(0);
  int j = get_global_id(1);

@@

\lstset{style=customcl,
        caption={Laplace OpenCL Filter Vectorized with Synthesized Loads.},
        breaklines=true,
        breakautoindent=false,
        label={lst:laplace_opencl_filter_vector_synth}}
\begin{lstlisting}
  uchar16 v11_ = vload16( 0, src + 3*width*(j-1) + 3*5*i - 3 );
  uchar16 v13_ = vload16( 0, src + 3*width*(j-1) + 3*5*i + 3 );
  uchar16 v12_ = uchar16( v11_.s3456789a, v13_.s56789abc );
  uchar16 v21_ = vload16( 0, src + 3*width*(j  ) + 3*5*i - 3 );
  uchar16 v23_ = vload16( 0, src + 3*width*(j  ) + 3*5*i + 3 );
  uchar16 v22_ = uchar16( v21_.s3456789a, v23_.s56789abc );
  uchar16 v31_ = vload16( 0, src + 3*width*(j+1) + 3*5*i - 3 );
  uchar16 v33_ = vload16( 0, src + 3*width*(j+1) + 3*5*i + 3 );
  uchar16 v32_ = uchar16( v31_.s3456789a, v33_.s56789abc );
\end{lstlisting}

@@

\lstset{style=customcl,
        caption={Laplace OpenCL Filter Vectorized Using \textit{short}.},
        breaklines=true,
        breakautoindent=false,
        label={lst:laplace_opencl_filter_short}}
\begin{lstlisting}
  short16 v11 = convert_short16(v11_);
  short16 v12 = convert_short16(v12_);
  short16 v13 = convert_short16(v13_);
  short16 v21 = convert_short16(v21_);
  short16 v22 = convert_short16(v22_);
  short16 v23 = convert_short16(v23_);
  short16 v31 = convert_short16(v31_);
  short16 v32 = convert_short16(v32_);
  short16 v33 = convert_short16(v33_);

@@

Nonetheless, the process is frustrating. The optimizations are never evaluated
independently from one another. Some were arbitrarily configured (like the
number of pixels chosen for instance). Testing the different combinations of
those optimizations and the different parameters would be very costly in
developer time and, moreover, like all repetitive and tedious tasks, error
prone. Thus every created kernel would have to be thoroughly tested to ensure
no error was done.

@@

In order to clearly differentiate what is going to be generated from what is
related to manipulations in the host language four keywords were defined. They
are \emph{decl}, \emph{pr}, \emph{opn} and \emph{close}. As the language is an
Embedded Domain Specific Language (EDSL), these four keywords are methods in
the BOAST namespace. Sample usage of these keywords will be found in the next
Figures.
%Usage of these keywords can be found on Figure~\ref{fig:keywords}.

@@

\begin{figure}
\lstset{style=BOAST,
        caption={BOAST code.}, 
        label={lst:BOAST_algebra},
        numbers=left }
\begin{lstlisting}
i = BOAST::Int("i") # or BOAST::Variable::new("i", Int)
k = BOAST::Int("k", :size => 8)
l = BOAST::Real("l", :dim => [ BOAST::Dim(-5, 21) ], :local => true )
BOAST::decl i, k, l
BOAST::pr i === 5
j = i + 5
BOAST::pr k === j * 2
BOAST::pr l[k] === 1.0
BOAST::register_funccall("sin")
BOAST::pr l[k+1] === BOAST::sin(j)
\end{lstlisting}

@@

\begin{figure}
\lstset{style=BOAST,
        caption={BOAST code.}, 
        label={lst:BOAST_control},
        numbers=left }
\begin{lstlisting}
BOAST::register_funccall("modulo")
i = BOAST::Int("i")
j = BOAST::Int("j")
BOAST::pr j === 0
BOAST::pr BOAST::For( i, 0, 100 ) {
  BOAST::pr BOAST::If( BOAST::modulo(i,7) == 0 ) {
    BOAST::pr j === j + 1
  }
}
\end{lstlisting}

@@

\begin{figure}
\lstset{style=BOAST,
        caption={BOAST code.}, 
        label={lst:BOAST_procedure},
        numbers=left }
\begin{lstlisting}
n = Int("n", :dir => :in)
ndat = Int("ndat", :dir => :in)
x = Real("x", :dir => :in, :dim => [ Dim(0, n-1), Dim(ndat) ])
y = Real("y", :dir => :out, :dim => [ Dim(ndat), Dim(0, n-1) ])
p = Procedure("magicfilter", [n, ndat, x, y])
opn p
close p
\end{lstlisting}
